Analyzing the EDHOC Key Establishment Protocol using TAMARIN

### 2022-07-10 Journal Version/journal-version.tex
Added files with the new text

\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
\author{
        Karl Norrman and 
        Vaishnavi Sundararajan and
        Alessandro Bruni
}

\begin{document}
\maketitle
%
\begin{abstract}
Constrained IoT devices are becoming ubiquitous in society
and there is a need for secure communication protocols that respect the
constraints under which these devices operate.
%
\mEdhoc{} is an authenticated key establishment protocol for constrained IoT
devices, currently being standardized by the Internet Engineering Task
Force (IETF).
%
A rudimentary version of \mEdhoc{} with only two key establishment methods was
formally analyzed in 2018.
%
Since then, the protocol has evolved significantly and several new key
establishment methods have been added.
%
In this paper, we present a formal analysis of all \mEdhoc{} methods in an
enhanced symbolic Dolev-Yao model using the \mTamarin{} tool.
%
We show that not all methods satisfy the authentication notion
injective agreement, but that all do satisfy a notion of implicit
authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
%
We identify other weaknesses to which we propose improvements
%
For example, a party may intend to establish a session key with a certain peer,
but ending up establishing it with another, trusted but compromised, peer.
%
We communicated our findings and proposals to the IETF, which
has incorporated some of these in newer versions of the standard.
\end{abstract}
%
%-------------------------------------------------------------------------- sec
\section{\uppercase{Introduction}}
\label{sec:introduction}

As IoT devices become more prevalent and get involved in progressively sensitive
functions in society, the need to secure their communications
becomes increasingly important.
%
Most security analysis has focused on computationally
strong devices, such as cars and web-cameras, where existing protocols like
\mDandTls{} suffice.
%
Constrained devices, on the other hand, which operate under severe
bandwidth and energy consumption restrictions, have received much less
attention.
%
These devices may be simple sensors, which only relay environment
measurements to a server every hour, but need to function autonomously without
maintenance for long periods of time.
%
The IETF standardized the Object Security for
Constrained RESTful Environments (\mOscore{}) protocol to secure communications
between constrained devices~\cite{rfc8613}.
%
However, the \mOscore{} protocol requires a pre-established security context.
%
The IETF has been discussing requirements and mechanisms for a key
exchange protocol, named Ephemeral Diffie-Hellman Over COSE (\mEdhoc), for
establishing \mOscore{} security contexts.
%
Naturally, \mEdhoc{} must work under the same constrained requirements as
\mOscore{} itself.
%
While not all use cases for \mEdhoc{} are firmly set, the overall goal is to
establish an \mOscore{} security context, under message size limitations.
%
It is therefore important to ensure that \mEdhoc{} satisfies fundamental
security properties expected from a key exchange protocol.
%The work is now being done in the IETF Lightweight Authenticated Key Exchange
%(LAKE) working group.
%

The first incarnation of \mEdhoc{} appeared in March 2016.
%
It contained two different key establishment methods, one based on a
pre-shared Diffie-Hellman (DH) cryptographic core%
\footnote{By a \emph{cryptographic core}, or simply core, we mean an academic protocol,
without encodings or application specific details required by an industrial
protocol.
%
A \emph{key establishment method} is a core with some such details added.
}
and a second based on a
variant of challenge-response signatures in the style of
\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
%
\mEdhoc{} is therefore a framework of several key establishment methods.
%
In May 2018, the core based on challenge-response signatures was replaced by
one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}.
%
Since then the protocol has undergone significant changes.
%
Three new cores, mixing challenge-response signatures and regular signature for
authentication, were added~\cite{our-analysis-selander-lake-edhoc-00}.
%

We formulate and formalize a security model covering all four key
establishment methods, which is important especially since the
\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} lacks a clear description
of the intended security model and overall security goals.
%

We perform the analysis in a symbolic Dolev-Yao model.
In this framework, we model messages as terms in an algebra, with operations such
as encryption, modelled as functions on these terms.
These functions are assumed perfect, e.g., one cannot decrypt an encrypted
message without access to the key.
The adversary, while unable to break encryption or reverse hashing, is modelled
as the network.
That is, the adversary, can block reroute, replay and modify messages at will.
A symbolic model like this, while slightly severe an abstraction, still allows
us to analyze \mEdhoc{} for logical flaws without incurring the complexity of
a computational model.
%
The standardization process is ongoing, with the authors releasing newer
versions of the \mSpec{} (see Section~\ref{sec:newdrafts} for more detail
about how these versions differ from the one analyzed here).
%

%-------------------------------------------------------------------------- sub
\subsection{Contributions}
\label{sec:contributions}
In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
establishment methods) using the \mTamarin{} tool~\cite{DBLP:conf/cav/MeierSCB13}.
%
We present a formal model we constructed of the protocol as given in the
\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
%

We give an explicit adversary model for the protocol and verify
properties such as session key material and entity authentication, and perfect
forward secrecy, for all four methods.
%

The model itself is valuable as a basis for verifying further updates in the
ongoing standardization.
%
It is publicly available~\cite{edhocTamarinRepo}.
%
It took several person-months to interpret the
specification and construct the model.
%
Termination requires a hand-crafted proof oracle to guide \mTamarin{}.
%

We show that not all \mEdhoc{}'s key establishment
methods provide authentication according to the injective agreement definition
on the session key material, and none on the initiator's identity.
%
However, we show that all methods fulfill an implicit agreement property
covering the session key material and the initiator's identity.
%
We identify a number of subtleties, ambiguities and weaknesses in the
specification.
%
For example, the authentication policy requirements allows situations where a
party establishes session key material with a trusted but compromised peer, even
though the intention was to establish it with a different trusted party.
%
We provide remedies for the identified issues and have
communicated these to the IETF and the specification authors, who incorporated
some of our suggestions and currently consider how to deal with the remaining
ones.
%

%-------------------------------------------------------------------------- sub
\subsection{Comparison with Related Work}
The May 2018 version of \mEdhoc{} was formally analyzed by
\cite{DBLP:conf/secsr/BruniJPS18} using the \mProverif{}
tool~\cite{DBLP:conf/csfw/Blanchet01}.
%
Their analysis covered a pre-shared key authenticated core and one
based on \mSigma.
%
The properties checked for therein were secrecy, PFS and integrity of
application data, identity protection against an active adversary,
and strong authentication.
%

In contrast to the key establishment methods analyzed by Bruni et~al., which
were based on the well-understood pre-shared key DH and \mSigma{} protocols,
the three newly added
methods combine two unilateral authentication protocols with the goal to
constructing mutual authentication protocols.
%
Combining two protocols, which individually provide unilateral authentication,
is not guaranteed to result in a secure mutual authentication
protocol~\cite{DBLP:conf/ccs/Krawczyk16}.
%
Consequently, even though the framework is similar to the one analyzed by Bruni
et~al., the cryptographic underpinnings have significantly increased in
complexity, and is using mechanisms which have not previously been formally analyzed.
%
The set of properties we check for is also different.
%
Our analysis is further carried out using a different tool,
namely \mTamarin; different kinds of strategies to formulate and
successfully analyze the protocol are required when working with this tool.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{The \mEdhoc{} Protocol}}
\label{sec:edhoc}
\input{edhocProtocol}

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
%\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
The core \mEdhoc{} specification has limited functionality, few negotiable
parameters and processing rules.
%
Hence, we argue in this section that its attack surface
is limited from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined or
specified as (optional) extensions.
%
It is therefore difficult to draw the line between what is \mEdhoc{} and what is
an extension.
%
For example, the core specification does not include a
message transport mechanism.
%
This increases the risk of applications vulnerable to DoS via \mEdhoc{}, even
though \mEdhoc{} itself is arguably not particularly weak in this respect.
%

\paragraph{DoS mitigation is left to developers.}
Section 8.7 of the \mEdhoc{} specification\footnote{draft-ietf-lake-edhoc-14}
explains the design philosophy w.r.t.
DoS in two paragraphs and gives two recommendations.
%
The first being that \mEdhoc{} relies on lower layers to mitigate DoS, and an
example of DoS countermeasures is given: checking return routeability when
receiving the first message.
%
The second being that an application may try to determine (in some way) that an
apparently valid message is in fact probably a forgery and should be ignored or
cause processing to stop.
%

No recommendation is given on how to minimize DoS effects stemming from use of
error messages, which can be used in an arbitrary way in \mEdhoc{}.
%
While there is some work on DDoS aspects of \mCoap{}, and much on IoT in
general, this does not capture the particular consequences originating with
\mEdhoc{}.
%
We now elaborate on some considerations application developers needs to take
into account.
%

\paragraph{Amplification and role asymmetry.}
For the same reasons as given in the IKEv2 DDoS analysis
by~\cite{rfc8019} the responder is more vulnerable than the
initiator.
%
The key similarity between IKEv2 and \mEdhoc{} here is that the initiator is
anonymous in the first message and that creating the first message is
computationally cheap and requires no storage.
%
No element in the first \mEdhoc{} message can be verified by the receiver as
being legitimate as long as the format is correct and elliptic curve points
lie on the proposed curve.
%
\mEdhoc{} leave it to the application and its
transport layer to resolve any issues.
%
In addition to the computational resources the responder spends, \mEdhoc{}
execution may require the responder to
obtain and verify certificate chains, and verify
certificate-revocation servers as a result of the spoofed initial message.
%
For example, the content of the \mADone{} information element may hint the
responder to take further actions.
%

The asymmetry between the adversary's effort and the responder's may hence be
significant.
%
Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
be considered even more carefully than for IKEv2~\cite{rfc8019}.
%

Depending on the use case, the first \mEdhoc{} message may be sent by either
party in a session.
%
For example, even though it may seem intuitive that an electronic
key-card initiates the connection towards lock, the first \mEdhoc{} message
could be included in the response from the lock.
%
This means the lock can postpone heavy verification actions until it has
authenticated the key card after receiving the third \mEdhoc{} message.
%

\paragraph{Reflection Attacks.}
Reflection attacks are attacks where an adversary spoofs the source address of a
victim in a message and thereby tricks the responder to send a message to the
victim.
%
\mEdhoc{} leaves protection against this to the transport protocol and
application, and suggests using a return path reachability method similar to IKEv2.
%
We note that such reachability mechanisms can easily be used to make a responder
send reachability requests to arbitrary addressable targets.
%
Depending on the design of such reachability mechanisms, they may require that
unsolicited reachability requests are discarded, limiting the effects of
anonymous DoS attacks like these.
%

In situations where the responder does not need to perform additional
communications to verify certificates or similar, a reachability mechanism may
be more expensive in terms of time and storage compared to continue with the
protocol and keeping a half-open state for a period of time.
%
Doing so would also avoid creating secondary DoS effects on certificate storage
nodes or certificate revocation servers.
%
The latter may cause a systems-level issue if many distributed
\mEdhoc{} responders frequently requests such servers.
%
However, if round-trip times are large, and half-open states hence are kept
by the receiver for a long time, a reachability mechanism can trade
communication for storage.
\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
up a simple model and derive an expression for trade-offs? Could be useful for
implementors to consider, but low academic value. This would capture also
amplification.}
%

%\mPoint{In a constrained device randomness may be easily depleted}
%Assume an adversary generate (deterministic) bogus message 1s and send these at
%a sufficiently high rate to the responder.
%%
%\mEdhoc{} specifically targets constrained devices, that presumably may not
%always have good randomness sources.
%%
%Is it possible to deplete or significantly reduce the entropy of the source and
%in this way be able to predict the responder's random input to other sessions
%(with other initiators)?
%%
%The LAKE WG have been talking about effects of poor randomness on signatures, so
%this should be an issue, no?
%%

\mPoint{\mEdhoc{} can deadlock if not used correctly}

An adversary may send forged error messages, and specifically reject proposed
ciphersuites.
%
The initiator is recommended to not try that ciphersuite with the responder
again in a new session for a while.
%
If the adversary rejects all options for a initiator this way, it can prevent
communication between the initiator and that responder until the cache is
cleared.
%
Potentially, an adversary could lock a party of a group out of communication all
together this way.
%
This needs to be considered by applications.
%

While we have not conducted a formal analysis of the fact, the structure of all
methods are so similar that it seems plausible that none of them would be
stronger against DDoS than the others.
%

Implementers are well advised to consult \cite{rfc8019}, which provides much
good general information about DoS for key establishment
protocols with the 3-message structure employed by \mEdhoc{}.
%
However, ultimately, a product-specific risk assessment is required to
determine appropriate DDoS mitigations.
%

\section{\uppercase{Discussion}}
\label{sec:discussion}
There are a few places where \mEdhoc{} can be improved,
which we found during this work and communicated to the authors.
%
We discuss them below.
%

>>>>

%\title{Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices}
\title{Analyzing the EDHOC Key Establishment Protocol using TAMARIN}
\author{
        Karl Norrman and 
        Vaishnavi Sundararajan and
        Alessandro Bruni
}

\begin{document}
\maketitle
%
\begin{abstract}
Given how common IoT devices that use constrained resources are becoming today, the need of the hour is communication protocols which can operate securely under such limitations.
%
For a few years, the Internet Engineering Task Force (IETF) has been working to standardize \mEdhoc{}, an authenticated key establishment protocol for such constrained IoT devices.
%
The first version of \mEdhoc{} was proposed in 2016.
%
In 2018, Bruni et al~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.  
%
%By now, the protocol has been fleshed out much more, and has been augmented with multiple new key establishment methods.
By 2021, the protocol had been fleshed out much more, with multiple new key establishment methods, and this version was formally analyzed using the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
%
Here, we build on this work, and use \mTamarin{} to formally analyze the key establishment methods in the current version of \mEdhoc{}, as well as discuss some ramifications of the choices made while designing the protocol.
\vnote{We need to see what more we can say here, about the new stuff we're potentially adding}
%
%We show that not all methods satisfy the authentication notion
%injective agreement, but that all do satisfy a notion of implicit
%authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
%%
%We identify other weaknesses to which we propose improvements
%%
%For example, a party may intend to establish a session key with a certain peer,
%but ending up establishing it with another, trusted but compromised, peer.
%%
%We communicated our findings and proposals to the IETF, which
%has incorporated some of these in newer versions of the standard.
%Constrained IoT devices are becoming ubiquitous in society
%and there is a need for secure communication protocols that respect the
%constraints under which these devices operate.
%%
%\mEdhoc{} is an authenticated key establishment protocol for constrained IoT
%devices, currently being standardized by the Internet Engineering Task
%Force (IETF).
%%
%A rudimentary version of \mEdhoc{} with only two key establishment methods was
%formally analyzed in 2018.
%%
%Since then, the protocol has evolved significantly and several new key
%establishment methods have been added.
%%
%In this paper, we present a formal analysis of all \mEdhoc{} methods in an
%enhanced symbolic Dolev-Yao model using the \mTamarin{} tool.
%%
%We show that not all methods satisfy the authentication notion
%injective agreement, but that all do satisfy a notion of implicit
%authentication, as well as Perfect Forward Secrecy (PFS) of the session key material.
%%
%We identify other weaknesses to which we propose improvements
%%
%For example, a party may intend to establish a session key with a certain peer,
%but ending up establishing it with another, trusted but compromised, peer.
%%
%We communicated our findings and proposals to the IETF, which
%has incorporated some of these in newer versions of the standard.
\end{abstract}
%
%-------------------------------------------------------------------------- sec
\section{\uppercase{Introduction}}
\label{sec:introduction}
%%%%%%% START HERE %%%%%%%%%%
\input{intro}

%As IoT devices become more prevalent and get involved in progressively sensitive
%functions in society, the need to secure their communications
%becomes increasingly important.
%%
%Most security analysis has focused on computationally
%strong devices, such as cars and web-cameras, where existing protocols like
%\mDandTls{} suffice.
%%
%Constrained devices, on the other hand, which operate under severe
%bandwidth and energy consumption restrictions, have received much less
%attention.
%%
%These devices may be simple sensors, which only relay environment
%measurements to a server every hour, but need to function autonomously without
%maintenance for long periods of time.
%%
%The IETF standardized the Object Security for
%Constrained RESTful Environments (\mOscore{}) protocol to secure communications
%between constrained devices~\cite{rfc8613}.
%%
%However, the \mOscore{} protocol requires a pre-established security context.
%%
%The IETF has been discussing requirements and mechanisms for a key
%exchange protocol, named Ephemeral Diffie-Hellman Over COSE (\mEdhoc), for
%establishing \mOscore{} security contexts.
%%
%Naturally, \mEdhoc{} must work under the same constrained requirements as
%\mOscore{} itself.
%%
%While not all use cases for \mEdhoc{} are firmly set, the overall goal is to
%establish an \mOscore{} security context, under message size limitations.
%%
%It is therefore important to ensure that \mEdhoc{} satisfies fundamental
%security properties expected from a key exchange protocol.
%%The work is now being done in the IETF Lightweight Authenticated Key Exchange
%%(LAKE) working group.
%%
%
%The first incarnation of \mEdhoc{} appeared in March 2016.
%%
%It contained two different key establishment methods, one based on a
%pre-shared Diffie-Hellman (DH) cryptographic core%
%\footnote{By a \emph{cryptographic core}, or simply core, we mean an academic protocol,
%without encodings or application specific details required by an industrial
%protocol.
%%
%A \emph{key establishment method} is a core with some such details added.
%}
%and a second based on a
%variant of challenge-response signatures in the style of
%\mOptls{}~\cite{DBLP:conf/eurosp/KrawczykW16}.
%%
%\mEdhoc{} is therefore a framework of several key establishment methods.
%%
%In May 2018, the core based on challenge-response signatures was replaced by
%one based on \mSigma{} (SIGn-and-MAc)~\cite{sigma,bruni-analysis-selander-ace-cose-ecdhe-08}.
%%
%Since then the protocol has undergone significant changes.
%%
%Three new cores, mixing challenge-response signatures and regular signature for
%authentication, were added~\cite{our-analysis-selander-lake-edhoc-00}.
%%
%
%We formulate and formalize a security model covering all four key
%establishment methods, which is important especially since the
%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00} lacks a clear description
%of the intended security model and overall security goals.
%%
%
%We perform the analysis in a symbolic Dolev-Yao model.
%In this framework, we model messages as terms in an algebra, with operations such
%as encryption, modelled as functions on these terms.
%These functions are assumed perfect, e.g., one cannot decrypt an encrypted
%message without access to the key.
%The adversary, while unable to break encryption or reverse hashing, is modelled
%as the network.
%That is, the adversary, can block reroute, replay and modify messages at will.
%A symbolic model like this, while slightly severe an abstraction, still allows
%us to analyze \mEdhoc{} for logical flaws without incurring the complexity of
%a computational model.
%%
%The standardization process is ongoing, with the authors releasing newer
%versions of the \mSpec{} (see Section~\ref{sec:newdrafts} for more detail
%about how these versions differ from the one analyzed here).
%%
%
%%-------------------------------------------------------------------------- sub
%\subsection{Contributions}
%\label{sec:contributions}
%In this paper, we formally analyze the \mEdhoc{} protocol (with its four key
%establishment methods) using the \mTamarin{} tool~\cite{DBLP:conf/cav/MeierSCB13}.
%%
%We present a formal model we constructed of the protocol as given in the
%\mSpec{}~\cite{our-analysis-selander-lake-edhoc-00}.
%%
%
%We give an explicit adversary model for the protocol and verify
%properties such as session key material and entity authentication, and perfect
%forward secrecy, for all four methods.
%%
%
%The model itself is valuable as a basis for verifying further updates in the
%ongoing standardization.
%%
%It is publicly available~\cite{edhocTamarinRepo}.
%%
%It took several person-months to interpret the
%specification and construct the model.
%%
%Termination requires a hand-crafted proof oracle to guide \mTamarin{}.
%%
%
%We show that not all \mEdhoc{}'s key establishment
%methods provide authentication according to the injective agreement definition
%on the session key material, and none on the initiator's identity.
%%
%However, we show that all methods fulfill an implicit agreement property
%covering the session key material and the initiator's identity.
%%
%We identify a number of subtleties, ambiguities and weaknesses in the
%specification.
%%
%For example, the authentication policy requirements allows situations where a
%party establishes session key material with a trusted but compromised peer, even
%though the intention was to establish it with a different trusted party.
%%
%We provide remedies for the identified issues and have
%communicated these to the IETF and the specification authors, who incorporated
%some of our suggestions and currently consider how to deal with the remaining
%ones.
%%
%
%%-------------------------------------------------------------------------- sub
%\subsection{Comparison with Related Work}
%The May 2018 version of \mEdhoc{} was formally analyzed by
%\cite{DBLP:conf/secsr/BruniJPS18} using the \mProverif{}
%tool~\cite{DBLP:conf/csfw/Blanchet01}.
%%
%Their analysis covered a pre-shared key authenticated core and one
%based on \mSigma.
%%
%The properties checked for therein were secrecy, PFS and integrity of
%application data, identity protection against an active adversary,
%and strong authentication.
%%
%
%In contrast to the key establishment methods analyzed by Bruni et~al., which
%were based on the well-understood pre-shared key DH and \mSigma{} protocols,
%the three newly added
%methods combine two unilateral authentication protocols with the goal to
%constructing mutual authentication protocols.
%%
%Combining two protocols, which individually provide unilateral authentication,
%is not guaranteed to result in a secure mutual authentication
%protocol~\cite{DBLP:conf/ccs/Krawczyk16}.
%%
%Consequently, even though the framework is similar to the one analyzed by Bruni
%et~al., the cryptographic underpinnings have significantly increased in
%complexity, and is using mechanisms which have not previously been formally analyzed.
%%
%The set of properties we check for is also different.
%%
%Our analysis is further carried out using a different tool,
%namely \mTamarin; different kinds of strategies to formulate and
%successfully analyze the protocol are required when working with this tool.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{The \mEdhoc{} Protocol}}
\label{sec:edhoc}
\input{protocol}

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
\mEdhoc{} itself has limited functionality, few negotiable parameters and
processing rules
%
Hence, we argue in this section that the attack surface
is small from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined.
%
This increases the risk of applications vulnerable to DoS even if \mEdhoc{} is
arguably not vulnerable to DoS itself.
%

\mPoint{\mEdhoc{} gives a little info of questionable use for developers}
Section 8.7 of the \mEdhoc{} specification explains the design philosophy w.r.t.
DoS in two paragraphs and gives two recommendations.
%
The first being that mitigation of DoS based on sending proper messages should
be handled by lower layers, e.g., by checking reachability of the sending
party\footnote{Karl: this may be another way to perform reflection DoS attacks}.
%
The second being that an application may try to determine (in some way) that an
apparently valid message is in fact probably a forgery and should be ignored or
cause processing to stop.
%

\mPoint{Responder is more vulnerable than initiator}.
For the same reasons as given in the IKEv2 DDoS analysis
by Nir and Smyslov~\cite{rfc8019} the responder is more vulnerable than the
initiator.
%
From a roles and message passing perspective, IKEv2 is very similar to all
methods in \mEdhoc{}.
%
The RFC provides much good advice for protocols with this structure that
\mEdhoc{} using applications can reuse.
%
We should look more into this.
%

\mPoint{Amplification}
An adversary can fill a message 1 with easily generated bogus data and send this
to a victim responder.
%
There is little work for the initiator, but both symmetric and asymmetric key
operations required by the responder.
%
\mEdhoc{} does not protect against this, but refer to the application and its
transport layer to resolve the problem.
%
This is a design decision, see Section 8.7 of \mEdhoc{} specification.
%
In addition to the computational resources the responder spends, it may also
need to obtain certificate chains and verify certificate revocation servers as a
result of the spoofed message.
%
The asymmetry between the adversaries effort and the responder may hence be
significant.
%
Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
be considered even more carefully than for IKEv2~\cite{rfc8019}.
%

Recommend that the role of responder is chosen carefully.
%
The responder is most vulnerable.
%
For example, is it better that a lock or one of the key cards takes on the
responder role and becomes vulnerable?
%

\mPoint{Reflection Attacks}
Reflection attacks are attacks where an adversary spoofs the source address of a
victim in a message and thereby tricks the responder to send a message to the
victim.
%
\mEdhoc{} leaves protection against this to the transport protocol and
application, and suggests using a path reachability method similar to IKEv2.
%
In situations where the responder does not need to perform additional
communications to verify certificates or similar, a reachability mechanism may
be more expensive than performing the computations.
\footnote{Find ref of that computation is often cheaper than radio
transmission, at least energy-wise.}
%
However, if round-trip times are large, and half-open states hence are kept
by the receiver for a long time, a reachability mechanism can trade
communication for storage.
\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
up a simple model and derive an expression for trade-offs? Could be useful for
implementors to consider, but low academic value. This would capture also
amplification.}
%

\mPoint{In a constrained device randomness may be easily depleted}
Assume an adversary generate (deterministic) bogus message 1s and send these at
a sufficiently high rate to the responder.
%
\mEdhoc{} specifically targets constrained devices, that presumably may not
always have good randomness sources.
%
Is it possible to deplete or significantly reduce the entropy of the source and
in this way be able to predict the responder's random input to other sessions
(with other initiators)?
%
The LAKE WG have been talking about effects of poor randomness on signatures, so
this should be an issue, no?
%

\mPoint{\mEdhoc{} can deadlock if not used correctly}
An adversary may send forged error messages, and specifically reject proposed
ciphersuites.
%
The initiator is recommended to not try that ciphersuite with the responder
again in a new session for a while.
%
If the adversary rejects all options for a initiator this way, it can prevent
communication between the initiator and that responder until the cache is
cleared.
%
Potentially, an adversary could lock a party of a group out of communication all
together this way.
%
This needs to be considered by applications.
%

\mPoint{Is any of the methods better than the others from a DoS perspective?}
We considered this for quite some time, but there seems to be little difference
in this regard.
%

%-------------------------------------------------------------------------- sec
\section{\uppercase{Discussion}}
\label{sec:discussion}
There are a few places where \mEdhoc{} can be improved,
which we found during this work and communicated to the authors.
%
We discuss them below.
%
